Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-07-25 Grassmannian_cells.tex
Added definitions of piecewise compatibility to section 6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
For $m\ge1$, let $D_m$ denote the maximal Dyck path in the lattice rectangle with corner vertices $(0,0)$ and $(u_m,u_{m-1})$.  
More precisely, $D_m$ is the lattice path which begins at $(0,0)$, takes East and North steps to end at $(u_m,u_{m-1})$, and never passes above the main diagonal joining $(0,0)$ and $(u_m,u_{m-1})$.  
It is maximal in the sense that any lattice point lying strictly above $D_m$ also lies above the main diagonal.  
Identify the edges of $D_m$ with the poset $E_m=\{1,\ldots,u_m+u_{m-1}\}$ where edges of $D_m$ are taken in the natural order beginning from $(0,0)$.
Given edges $e,e'\in E_m$ with $e<e'$, write $ee'$ for the smallest subpath of $D_m$ containing $e$ and $e'$, in particular $ee$ is the subpath containing the single edge $e$.

Let $H_m=\{h_1,\ldots,h_{u_m}\}$ and $V_m=\{v_1,\ldots,v_{u_{m-1}}\}$ denote the horizontal and vertical edges of $D_m$ respectively.
\begin{definition}
  A pair of subsets $S_H\subset H_m$ and $S_V\subset V_m$ is called compatible if: 
  for each pair $(h,v)\in S_H\times S_V$ with $h<v$ there exists an edge $e\in hv$ so that at least one of the following holds
  \[e\ne v\qquad\text{and}\qquad |he\cap S_V|=n|he\cap S_H|\]
  or
  \[e\ne h\qquad\text{and}\qquad |ev\cap S_H|=n|ev\cap S_V|.\]
  Write $\cC_m$ for the collection of all pairs $(S_H,S_V)$ which are compatible as above.
\end{definition}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
For $m\ge1$, let $D_m$ denote the maximal Dyck path in the lattice rectangle with corner vertices $(0,0)$ and $(u_m,u_{m-1})$.  
More precisely, $D_m$ is the lattice path which begins at $(0,0)$, takes East and North steps to end at $(u_m,u_{m-1})$, and never passes above the main diagonal joining $(0,0)$ and $(u_m,u_{m-1})$.  
It is maximal in the sense that any lattice point lying strictly above $D_m$ also lies above the main diagonal.

The maximal Dyck paths $D_m$, $m\ge1$, exhibit the following recursive structure.
In what follows we assume $n\ge2$.
\begin{theorem}
  \cite[Corollary 2.4]{rupel}
  \label{th:dyck path recursion}
  For $n\ge2$, the maximal Dyck path $D_m$, $m\ge1$, can be constructed recursively as follows:
  \begin{enumerate}
    \item $D_1$ consists of a single horizontal edge;
    \item $D_2$ consists of $n$ consecutive horizontal edges followed by a vertical edge;
    \item $D_m$, $m\ge3$, consists of $n-1$ copies of $D_{m-1}$ followed by a copy of $D_{m-1}$ with its first $D_{m-2}$ removed.
  \end{enumerate}
\end{theorem}

For $m\ge2$ and $1\le r\le n-1$, write $D_m^{(r)}$ for the maximal Dyck path obtained from $D_m$ by removing the first $r$ copies of $D_{m-1}$.
Note that by Theorem~\ref{th:dyck path recursion}, we have a natural identification of $D_m^{(n-1)}$, $m\ge3$, with $D_{m-1}^{(1)}$.
For fixed $m\ge2$ and $1\le r\le n-1$, we write $D_{m-1,r}$ for the $r$-th copy of $D_{m-1}$ inside $D_m$.
For notational convenience we also set $D_{m-1,n}:=D_m^{(n-1)}$  even though this maximal Dyck path does not identify with a copy of $D_{m-1}$.

We identify the edges of $D_m$ with the ordered set $E_m=\{1,\ldots,u_m+u_{m-1}\}$, where edges of $D_m$ are taken in the natural order beginning from $(0,0)$.
Let $E_m=H_m\sqcup V_m$, where $H_m=\{h_1,\ldots,h_{u_m}\}$ and $V_m=\{v_1,\ldots,v_{u_{m-1}}\}$ denote the horizontal and vertical edges of $D_m$ respectively.
Following Theorem~\ref{th:dyck path recursion}, we partition the edges as $E_m=\bigsqcup_{r=1}^n E_{m-1,r}$, where $E_{m-1,r}$ denotes the edges of $D_{m-1,r}$.
The set $E_{m-1,r}$ is naturally partitioned into its subsets $H_{m-1,r}$ and $V_{m-1,r}$ of horizontal and vertical edges.

Given edges $e,e'\in E_m$ with $e<e'$, write $ee'$ for the shortest subpath of $D_m$ containing $e$ and $e'$, in particular $ee$ is the subpath containing the single edge $e$.
\begin{definition}
  A pair of subsets $S_H\subset H_m$ and $S_V\subset V_m$ is called \emph{compatible} if: 
  for each pair $(h,v)\in S_H\times S_V$ with $h<v$ there exists an edge $e\in hv$ so that at least one of the following holds
  \begin{equation}
    \label{eq:hgc}
    e\ne v\qquad\text{and}\qquad |he\cap S_V|=n|he\cap S_H|
  \end{equation}
  or
  \begin{equation}
    \label{eq:vgc}
    e\ne h\qquad\text{and}\qquad |ev\cap S_H|=n|ev\cap S_V|.
  \end{equation}
  Write $\cC_m$ for the collection of all pairs $(S_H,S_V)$ which are compatible as above.
\end{definition}
\begin{remark}
  This notion of compatibility extends naturally to the maximal Dyck paths $D_m^{(r)}$, $1\le r\le n-1$.  Write $\cC_m^{(r)}$ for the set of all compatible pairs in $D_m^{(r)}$.
\end{remark}

The recursive structure of the maximal Dyck paths from Theorem~\ref{th:dyck path recursion} gives rise to a recursive characterization of compatible pairs.
\begin{definition}
  \cite[Definition 3.11]{rupel}
  A pair of subsets $S_H\subset H_m$ and $S_V\subset V_m$ is called \emph{piecewise compatible} if one of the conditions \eqref{eq:hgc} or \eqref{eq:vgc} is satisfied for each pair $(h,v)\in(S_H\cap H_{m-1,r})\times(S_V\cap V_{m-1,r})$ $h\in H_{m-1,r}$ and $v\in V_{m-1,r}$ for some $1\le r\le n-1$.
\end{definition}

@@

\bibitem{rupel} Rupel, D.: Rank 2 Non-Commutative Laurent Phenomenon and Pseudo-Positivity, arXiv.
